---
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---
# Exploratory Data Analysis

The dataset for this report analysis is **Hotel Booking Demand,** which may be accessed in the **Assignments for final reports** section of MS-Teams as well as on the Kaggle platform: https://www.kaggle.com/jessemostipak/hotel-booking-demand.

This data collection comprises booking information for a city hotel and a resort hotel, including when the booking was made, duration of stay, number of adults, children, and/or babies, and number of available parking spaces, among other things.

Three datasets will be used in the analysis:  
**Train Data**: About 60% of the units of the original dataset.  
**Validation Data**: About 20% of the units of the original dataset.  
**Test Data**: About 20% of the units of the original dataset.  
The Target Variable to be predicted:  **is_canceled** in the training dataset, which consists of two classes “0” or “1”. 0 means “NO” and 1 means “Yes”.

This dataset on hotel bookings can assist us in answering following queries:
Have you ever thought about when the optimal time to book a hotel stay is? Or what is the best length of stay to get the greatest daily rate? What if you wanted to know whether a hotel was likely to receive an unusually large number of special requests?
Also we can utilize this dataset to conduct research on a variety of issues such as booking cancellation prediction, customer segmentation, customer satiation, seasonality, and so on.

## Attribute Specification

In the start we need to import our train dataset for the exploratory data analysis.

```{r include=FALSE}
hb_train <- read.csv("hotel_bookings_train.csv")
```
```{r}
str(hb_train)
length(unique(hb_train$X))
```

The dataset contains initially 71633 observations of 32 variables. We will drop the first variable which is for the Id of the hotels which has all the unique values of 71633 in total.

```{r}
colSums(sapply(hb_train, is.na))
```

The dataset contain 2 missing values in the children variable so we will omit these missing values from the dataset.
When above transformation has been made, now the dataset has 31 variables with 71631 observations. Following are the variable type and description to understand the nature of the variable as per the information was collected.

**hotel:** Categorical, booking information for a city hotel or a resort hotel  
**is_canceled:** Categorical, Value indicating if the booking was canceled (1) or not (0)  
**lead_time:** Integer, Number of days that elapsed between the entering date of the booking into the PMS and the arrival date  
**arrival_date_year:** Integer,	Year of arrival date  
**arrival_date_month:** Categorical, Month of arrival date with 12 categories: “January” to “December”  
**arrival_date_week_number:**	Integer, Week number of the arrival date  
**arrival_date_day_of_month:** Integer,	Day of the month of the arrival date  
**stays_in_weekend_nights:** Integer,	Number of weekend nights (Saturday or Sunday) the guest stayed or booked to stay at the hotel  
**stays_in_week_nights:** Integer, Number of week nights (Monday to Friday) the guest stayed or booked to stay at the hotel  
**adults:** Integer, Number of adults  
**children:** Integer, Number of children  
**babies:** Integer, Number of babies  
**meal:** Categorical, Type of meal booked. Categories are presented in standard hospitality meal packages: Undefined/SC – no meal package; BB – Bed & Breakfast; HB – Half board (breakfast and one other meal – usually dinner);  
**country:** Categorical, Country of origin  
**market_segment:** Categorical, Market segment designation. In categories, the term “TA” means “Travel Agents” and “TO” means “Tour Operators”  
**distribution_channel:** Categorical, Booking distribution channel. The term “TA” means “Travel Agents” and “TO” means “Tour Operators”  
**is_repeated_guest:** Categorical, Value indicating if the booking name was from a repeated guest (1) or not (0)  
**previous_cancellations:** Integer, Number of previous bookings that were canceled by the customer prior to the current booking  
**agent:** Categorical, ID of the travel agency that made the booking  
**previous_bookings_not_canceled:** Integer, Number of previous bookings not canceled by the customer prior to the current booking  
**reserved_room_type:** Categorical, Code of room type reserved. Code is presented instead of designation for anonymity reasons  
**assigned_room_type:** Categorical, Code for the type of room assigned to the booking. Sometimes the assigned room type differs from the reserved room type due to hotel operation reasons (e.g. overbooking) or by customer request. Code is presented instead of designation for anonymity reasons  
**booking_changes:** Integer, Number of changes or amendments made to the booking from the moment the booking was entered on the PMS until the moment of check-in or cancellation  
**deposit_type** Categorical, Indication on if the customer made a deposit to guarantee the booking. This variable can assume three categories: No Deposit – no deposit was made; Non Refund – a deposit was made in the value of the total stay cost; Refundable – a deposit was made with a value under the total cost of stay  
**company** Categorical, ID of the company/entity that made the booking or responsible for paying the booking. ID is presented instead of designation for anonymity reasons  
**days_in_waiting_list:** Integer, Number of days the booking was in the waiting list before it was confirmed to the customer  
**customer_type:** Categorical, Type of booking, assuming one of four categories: Contract - when the booking has an allotment or other type of contract associated to it; Group – when the booking is associated to a group; Transient – when the booking is not part of a group or contract, and is not associated to other transient booking; Transient-party – when the booking is transient, but is associated to at least other transient booking
**adr:** Numeric,	Average Daily Rate as defined  
**required_car_parking_spaces:** Integer, Number of car parking spaces required by the customer  
**total_of_special_requests:** Integer, Number of special requests made by the customer (e.g. twin bed or high floor)  
**reservation_status_date:** Date, Date at which the last status was set. This variable can be used in conjunction with the Reservation Status to understand when was the booking canceled or when did the customer checked-out of the hotel

```{r include=FALSE}
hotel_booking <- na.omit(hb_train)
hotel_booking<- hotel_booking[-c(1)]
hotel_booking$hotel <- factor(hotel_booking$hotel)
hotel_booking$is_canceled <- factor(hotel_booking$is_canceled)
hotel_booking$arrival_date_month <- factor(hotel_booking$arrival_date_month)
hotel_booking$meal <- factor(hotel_booking$meal)
hotel_booking$market_segment <- factor(hotel_booking$market_segment)
hotel_booking$country <- factor(hotel_booking$country)
hotel_booking$distribution_channel <- factor(hotel_booking$distribution_channel)
hotel_booking$is_repeated_guest <- factor(hotel_booking$is_repeated_guest)
hotel_booking$agent <- factor(hotel_booking$agent)
hotel_booking$reserved_room_type <- factor(hotel_booking$reserved_room_type)
hotel_booking$assigned_room_type <- factor(hotel_booking$assigned_room_type)
hotel_booking$deposit_type <- factor(hotel_booking$deposit_type)
hotel_booking$company <- factor(hotel_booking$company)
hotel_booking$customer_type <- factor(hotel_booking$customer_type)
hotel_booking$reservation_status_date <- as.Date(hotel_booking$reservation_status_date)
```

```{r}
str(hotel_booking)
```

The Dataset consists of 14 variables that are categorical type, 16 variables that are of numeric type and 1 variable of Date type.
The main objective is to find out which criteria are the most successful in the forecast of cancellations of bookings. In this report, we will also classify our dataset by various techniques so that we can calculate the maximum precision for the main purpose. This helps us lower the booking cancellation after realizing which aspects have to be increased or decreased to mitigate the booking cancellation.

## Data Exploration

Given Below are the descriptive statistics of the hotel booking train dataset;

```{r echo=FALSE}
summary(hotel_booking)
```

We can see that in our dataset, certain variables have a mean that is higher than the median, while others have a median that is higher than the mean. So, in general, but not always, if the median is lower than the mean, we may see significant outliers at the high end of the distribution, while if the mean is lower than the median, we may see major outliers at the low end. To have a detailed overview of the variables, we will perform a univariate analysis to narrates the pattern of response to the variables.


```{r}
library(ggplot2)
ggplot(hotel_booking, aes(x = lead_time, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("lead_time") +
  ylab("density")
```

```{r}
ggplot(hotel_booking, aes(x = arrival_date_year, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("arrival_date_year") +
  ylab("density")
```

```{r}
ggplot(hotel_booking, aes(x = arrival_date_week_number, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("arrival_date_week_number") +
  ylab("density")
```
```{r}
ggplot(hotel_booking, aes(x = arrival_date_day_of_month, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("arrival_date_day_of_month") +
  ylab("density")
```
```{r}
ggplot(hotel_booking, aes(x = stays_in_weekend_nights, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("stays_in_weekend_nights") +
  ylab("density")
```

```{r}
ggplot(hotel_booking, aes(x = stays_in_week_nights, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("stays_in_week_nights") +
  ylab("density")
```

```{r}
ggplot(hotel_booking, aes(x = adults, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("adults") +
  ylab("density")
```
```{r}
ggplot(hotel_booking, aes(x = children, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("children") +
  ylab("density")
```

```{r}
ggplot(hotel_booking, aes(x = babies, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("babies") +
  ylab("density")
```
```{r}
ggplot(hotel_booking, aes(x = previous_cancellations, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("previous_cancellations") +
  ylab("density")
```

```{r}
ggplot(hotel_booking, aes(x = previous_bookings_not_canceled, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("previous_bookings_not_canceled") +
  ylab("density")
```

```{r}
ggplot(hotel_booking, aes(x = booking_changes, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("booking_changes") +
  ylab("density")
```

```{r}
ggplot(hotel_booking, aes(x = days_in_waiting_list, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("days_in_waiting_list") +
  ylab("density")
```

```{r}
ggplot(hotel_booking, aes(x = adr, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("adr") +
  ylab("density")
```

```{r}
ggplot(hotel_booking, aes(x = required_car_parking_spaces, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("required_car_parking_spaces") +
  ylab("density")
```

```{r}
ggplot(hotel_booking, aes(x = total_of_special_requests, y=..density..)) +
  geom_histogram(fill = "#6F8FAF", colour = "Black", bins = 5) +
  xlab("total_of_special_requests") +
  ylab("density")
```

```{r}
ggplot(data = hotel_booking,
       aes(
         x = hotel,
         y = prop.table(stat(count)),
         fill = is_canceled,
         label = scales::percent(prop.table(stat(count)))
       )) +
  geom_bar(position = position_dodge()) +
  geom_text(
    stat = "count",
    position = position_dodge(.9),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Reservation Status by Hotel Type", x = "Hotel Type", y = "Percentage") +
  theme_classic() +
  scale_fill_discrete(
    name = "Reservation Status",
    labels = c("Not Cancelled", "Cancelled")
  )
```

```{r}
ggplot(hotel_booking, aes(arrival_date_month, fill = is_canceled)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat = "count",
            aes(label = ..count..),
            position = position_dodge(.9),
    vjust = -0.5,
    size = 2) +
  scale_fill_discrete(
    name = "Reservation Status",
    label = c("Not Cancelled", "Cancelled")
  ) +
  labs(title = "Reservation Status by Month", x = "Month", y = "Count") +
  scale_x_discrete(labels = month.abb)
```

```{r}
ggplot(hotel_booking, aes(meal, fill = is_canceled)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat = "count",
            aes(label = ..count..),
            position = position_dodge(.9),
    vjust = -0.5,
    size = 3) +
  scale_fill_discrete(
    name = "Reservation Status",
    label = c("Not Cancelled", "Cancelled")
  ) +
  labs(title = "Reservation Status by Meal", x = "Meal", y = "Count")
```
```{r}
ggplot(hotel_booking, aes(country == "PRT" , fill = is_canceled)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat = "count",
            aes(label = ..count..),
            position = position_dodge(.9),
    vjust = -0.5,
    size = 3) +
  scale_fill_discrete(
    name = "Reservation Status",
    label = c("Not Cancelled", "Cancelled")
  ) +
  labs(title = "Reservation Status by Portugal and Other Countries", x = "Country Portugal", y = "Count")
```

```{r}
ggplot(hotel_booking, aes(deposit_type, fill = is_canceled)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat = "count",
            aes(label = ..count..),
            position = position_dodge(.9),
    vjust = -0.5,
    size = 3) +
  scale_fill_discrete(
    name = "Reservation Status",
    label = c("Not Cancelled", "Cancelled")
  ) +
  labs(title = "Reservation Status by Deposit type", x = "Deposit Type", y = "Count")
```

```{r}
ggplot(hotel_booking, aes(is_repeated_guest, fill = is_canceled)) +
  geom_bar(position = position_dodge()) +
    geom_text(stat = "count",
            aes(label = ..count..),
            position = position_dodge(.9),
    vjust = -0.5,
    size = 3) +
  labs(title = "Reservation Status by Repeated Guests", x = "Repeated Guest", y = "count") +
  theme_classic() +
  scale_fill_discrete(
    name = "Reservation Status",
    labels = c("Not Cancelled", "Cancelled")
  )
```

```{r}
ggplot(hotel_booking, aes(customer_type, fill = is_canceled)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat = "count",
            aes(label = ..count..),
            position = position_dodge(.9),
    vjust = -0.5,
    size = 3) +
  scale_fill_discrete(
    name = "Reservation Status",
    label = c("Not Cancelled", "Cancelled")
  ) +
  labs(title = "Reservation Status by Customer type", x = "Customers", y = "Count")
```

```{r}
ggplot(hotel_booking, aes(market_segment, fill = is_canceled)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat = "count",
            aes(label = ..count..),
            position = position_dodge(.9),
    vjust = -0.5,
    size = 3) +
  scale_fill_discrete(
    name = "Reservation Status",
    label = c("Not Cancelled", "Cancelled")
  ) +
  labs(title = "Reservation Status by Market Segment", x = "Market", y = "Count")
```

```{r}
ggplot(hotel_booking, aes(distribution_channel, fill = is_canceled)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat = "count",
            aes(label = ..count..),
            position = position_dodge(.9),
    vjust = -0.5,
    size = 3) +
  scale_fill_discrete(
    name = "Reservation Status",
    label = c("Not Cancelled", "Cancelled")
  ) +
  labs(title = "Reservation Status by Distribution Channel", x = "Distribution Channel", y = "Count")
```

```{r}
ggplot(hotel_booking, aes(company != "NULL", fill = is_canceled)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat = "count",
            aes(label = ..count..),
            position = position_dodge(.9),
    vjust = -0.5,
    size = 3) +
  scale_fill_discrete(
    name = "Reservation Status",
    label = c("Not Cancelled", "Cancelled")
  ) +
  labs(title = "Reservation Status by Company", x = "Company", y = "Count")
```

```{r}
ggplot(hotel_booking, aes(reserved_room_type, fill = is_canceled)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat = "count",
            aes(label = ..count..),
            position = position_dodge(.9),
    vjust = -0.5,
    size = 2) +
  scale_fill_discrete(
    name = "Reservation Status",
    label = c("Not Cancelled", "Cancelled")
  ) +
  labs(title = "Reservation Status by Reserved room type", x = "Reserved room type", y = "Count")
```

```{r}
ggplot(hotel_booking, aes(assigned_room_type, fill = is_canceled)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat = "count",
            aes(label = ..count..),
            position = position_dodge(.9),
    vjust = -0.5,
    size = 2) +
  scale_fill_discrete(
    name = "Reservation Status",
    label = c("Not Cancelled", "Cancelled")
  ) +
  labs(title = "Reservation Status by Assined room type", x = "Assigned room type", y = "Count")
```

```{r}
ggplot(hotel_booking, aes(agent == "NULL", fill = is_canceled)) +
  geom_bar(position = position_dodge()) +
  geom_text(stat = "count",
            aes(label = ..count..),
            position = position_dodge(.9),
    vjust = -0.5,
    size = 2) +
  scale_fill_discrete(
    name = "Reservation Status",
    label = c("Not Cancelled", "Cancelled")
  ) +
  labs(title = "Reservation Status by Agent", x = "Agent", y = "Count")
```

```{r}
str(hotel_booking)
```


```{r}
hb_sub <- hotel_booking[-c(1,2,5,13:17,20,21,23:25,27,31)]
library(psych)
## Warning: package 'psych' was built under R version 4.0.5
pairs.panels(hb_sub, main= "Correlation between the Variables",
ellipses = FALSE, gap = 0)
```

```{r}
library(corrplot)
corr.mat <- cor(hb_sub)
corrplot(corr.mat, method = "number")
```

```{r}
str(hotel_booking)
```


```{r}
library("Hmisc")
(res2 <- rcorr(as.matrix(hotel_booking[,-c(1,2,5,13:17,20,21,23:25,27,31)])))
```

