---
title: "R Notebook"
output: html_notebook
---

### train CSV

```{r}
train_hb <- read.csv("hotel_bookings_train.csv", header = T, sep = ",")[,-c(1)]
str(train_hb)
```
```{r}
train_hb <- na.omit(train_hb)
table(train_hb$hotel)
```
```{r}
train_hb$hotel[train_hb$hotel == "City Hotel"] <- 1
train_hb$hotel[train_hb$hotel == "Resort Hotel"] <- 2
table(train_hb$hotel)
```
```{r}
train_hb$hotel <- as.numeric(train_hb$hotel)
train_hb$arrival_date_month <-  factor(train_hb$arrival_date_month, levels = month.name)
table(train_hb$arrival_date_month)
```

```{r}
train_hb$arrival_date_month <- as.numeric(train_hb$arrival_date_month)
table(train_hb$arrival_date_month)
```

```{r}
table(train_hb$meal)
```

```{r}
train_hb$meal[train_hb$meal == "BB"] <- 1
train_hb$meal[train_hb$meal == "FB"] <- 2
train_hb$meal[train_hb$meal == "HB"] <- 3
train_hb$meal[train_hb$meal == "SC"] <- 4
train_hb$meal[train_hb$meal == "Undefined"] <- 5
table(train_hb$meal)
train_hb$meal <- as.numeric(train_hb$meal)
```

```{r}
table(train_hb$market_segment)
```

```{r}
train_hb$market_segment[train_hb$market_segment == "Aviation"] <- 1
train_hb$market_segment[train_hb$market_segment == "Complementary"] <- 2
train_hb$market_segment[train_hb$market_segment == "Corporate"] <- 3
train_hb$market_segment[train_hb$market_segment == "Direct"] <- 4
train_hb$market_segment[train_hb$market_segment == "Groups"] <- 5
train_hb$market_segment[train_hb$market_segment == "Offline TA/TO"] <- 6
train_hb$market_segment[train_hb$market_segment == "Online TA"] <- 7
table(train_hb$market_segment)
train_hb$market_segment <- as.numeric(train_hb$market_segment)
```

```{r}
table(train_hb$distribution_channel)
```
```{r}
train_hb$distribution_channel[train_hb$distribution_channel == "Corporate"] <- 1
train_hb$distribution_channel[train_hb$distribution_channel == "Direct"] <- 2
train_hb$distribution_channel[train_hb$distribution_channel == "GDS"] <- 3
train_hb$distribution_channel[train_hb$distribution_channel == "TA/TO"] <- 4
train_hb$distribution_channel[train_hb$distribution_channel == "Undefined"] <- 5
table(train_hb$distribution_channel)
train_hb$distribution_channel <- as.numeric(train_hb$distribution_channel)
```

```{r}
table(train_hb$is_repeated_guest)
train_hb$is_repeated_guest <- as.numeric(train_hb$is_repeated_guest)
table(train_hb$is_repeated_guest)
```

```{r}
table(train_hb$deposit_type)
```


```{r}
train_hb$deposit_type[train_hb$deposit_type == "No Deposit"] <- 1
train_hb$deposit_type[train_hb$deposit_type == "Non Refund"] <- 2
train_hb$deposit_type[train_hb$deposit_type == "Refundable"] <- 3
table(train_hb$deposit_type)
train_hb$deposit_type <- as.numeric(train_hb$deposit_type)
```

```{r}
table(train_hb$customer_type)
```


```{r}
train_hb$customer_type[train_hb$customer_type == "Contract"] <- 1
train_hb$customer_type[train_hb$customer_type == "Group"] <- 2
train_hb$customer_type[train_hb$customer_type == "Transient"] <- 3
train_hb$customer_type[train_hb$customer_type == "Transient-Party"] <- 4
table(train_hb$customer_type)
train_hb$customer_type <- as.numeric(train_hb$customer_type)
```

```{r}
str(train_hb)
```

### Validation CSV

```{r}
valid_hb <- read.csv("hotel_bookings_validation.csv", header = T, sep = ",")[,-c(1,15,21,22,25,26,32)]
str(valid_hb)
```


```{r}
valid_hb <- na.omit(valid_hb)
table(valid_hb$hotel)
```
```{r}
valid_hb$hotel[valid_hb$hotel == "City Hotel"] <- 1
valid_hb$hotel[valid_hb$hotel == "Resort Hotel"] <- 2
table(valid_hb$hotel)
```
```{r}
valid_hb$hotel <- as.numeric(valid_hb$hotel)
valid_hb$arrival_date_month <-  factor(valid_hb$arrival_date_month, levels = month.name)
table(valid_hb$arrival_date_month)
```

```{r}
valid_hb$arrival_date_month <- as.numeric(valid_hb$arrival_date_month)
table(valid_hb$arrival_date_month)
```

```{r}
table(valid_hb$meal)
```

```{r}
valid_hb$meal[valid_hb$meal == "BB"] <- 1
valid_hb$meal[valid_hb$meal == "FB"] <- 2
valid_hb$meal[valid_hb$meal == "HB"] <- 3
valid_hb$meal[valid_hb$meal == "SC"] <- 4
valid_hb$meal[valid_hb$meal == "Undefined"] <- 5
table(valid_hb$meal)
valid_hb$meal <- as.numeric(valid_hb$meal)
```

```{r}
table(valid_hb$market_segment)
```

```{r}
valid_hb$market_segment[valid_hb$market_segment == "Aviation"] <- 1
valid_hb$market_segment[valid_hb$market_segment == "Complementary"] <- 2
valid_hb$market_segment[valid_hb$market_segment == "Corporate"] <- 3
valid_hb$market_segment[valid_hb$market_segment == "Direct"] <- 4
valid_hb$market_segment[valid_hb$market_segment == "Groups"] <- 5
valid_hb$market_segment[valid_hb$market_segment == "Offline TA/TO"] <- 6
valid_hb$market_segment[valid_hb$market_segment == "Online TA"] <- 7
table(valid_hb$market_segment)
valid_hb$market_segment <- as.numeric(valid_hb$market_segment)
```

```{r}
table(valid_hb$distribution_channel)
```
```{r}
valid_hb$distribution_channel[valid_hb$distribution_channel == "Corporate"] <- 1
valid_hb$distribution_channel[valid_hb$distribution_channel == "Direct"] <- 2
valid_hb$distribution_channel[valid_hb$distribution_channel == "GDS"] <- 3
valid_hb$distribution_channel[valid_hb$distribution_channel == "TA/TO"] <- 4
valid_hb$distribution_channel[valid_hb$distribution_channel == "Undefined"] <- 5
table(valid_hb$distribution_channel)
valid_hb$distribution_channel <- as.numeric(valid_hb$distribution_channel)
```

```{r}
table(valid_hb$is_repeated_guest)
valid_hb$is_repeated_guest <- as.numeric(valid_hb$is_repeated_guest)
table(valid_hb$is_repeated_guest)
```

```{r}
table(valid_hb$deposit_type)
```


```{r}
valid_hb$deposit_type[valid_hb$deposit_type == "No Deposit"] <- 1
valid_hb$deposit_type[valid_hb$deposit_type == "Non Refund"] <- 2
valid_hb$deposit_type[valid_hb$deposit_type == "Refundable"] <- 3
table(valid_hb$deposit_type)
valid_hb$deposit_type <- as.numeric(valid_hb$deposit_type)
```

```{r}
table(valid_hb$customer_type)
```


```{r}
valid_hb$customer_type[valid_hb$customer_type == "Contract"] <- 1
valid_hb$customer_type[valid_hb$customer_type == "Group"] <- 2
valid_hb$customer_type[valid_hb$customer_type == "Transient"] <- 3
valid_hb$customer_type[valid_hb$customer_type == "Transient-Party"] <- 4
table(valid_hb$customer_type)
valid_hb$customer_type <- as.numeric(valid_hb$customer_type)
```

```{r}
str(valid_hb)
```


### Test CSV

```{r}
initial_test_hb <- read.csv("hotel_bookings_test.csv", header = T, sep = ",")[,-c(3)]
str(initial_test_hb)
```
```{r}
test_hb <- subset(initial_test_hb[,-c(1,14,20,21,24,25,31,32)])
```


```{r}
test_hb <- na.omit(test_hb)
table(test_hb$hotel)
```
```{r}
test_hb$hotel[test_hb$hotel == "City Hotel"] <- 1
test_hb$hotel[test_hb$hotel == "Resort Hotel"] <- 2
table(test_hb$hotel)
```
```{r}
test_hb$hotel <- as.numeric(test_hb$hotel)
test_hb$arrival_date_month <-  factor(test_hb$arrival_date_month, levels = month.name)
table(test_hb$arrival_date_month)
```

```{r}
test_hb$arrival_date_month <- as.numeric(test_hb$arrival_date_month)
table(test_hb$arrival_date_month)
```

```{r}
table(test_hb$meal)
```

```{r}
test_hb$meal[test_hb$meal == "BB"] <- 1
test_hb$meal[test_hb$meal == "FB"] <- 2
test_hb$meal[test_hb$meal == "HB"] <- 3
test_hb$meal[test_hb$meal == "SC"] <- 4
test_hb$meal[test_hb$meal == "Undefined"] <- 5
table(test_hb$meal)
test_hb$meal <- as.numeric(test_hb$meal)
```

```{r}
table(test_hb$market_segment)
```

```{r}
test_hb$market_segment[test_hb$market_segment == "Aviation"] <- 1
test_hb$market_segment[test_hb$market_segment == "Complementary"] <- 2
test_hb$market_segment[test_hb$market_segment == "Corporate"] <- 3
test_hb$market_segment[test_hb$market_segment == "Direct"] <- 4
test_hb$market_segment[test_hb$market_segment == "Groups"] <- 5
test_hb$market_segment[test_hb$market_segment == "Offline TA/TO"] <- 6
test_hb$market_segment[test_hb$market_segment == "Online TA"] <- 7
table(test_hb$market_segment)
test_hb$market_segment <- as.numeric(test_hb$market_segment)
```

```{r}
table(test_hb$distribution_channel)
```
```{r}
test_hb$distribution_channel[test_hb$distribution_channel == "Corporate"] <- 1
test_hb$distribution_channel[test_hb$distribution_channel == "Direct"] <- 2
test_hb$distribution_channel[test_hb$distribution_channel == "GDS"] <- 3
test_hb$distribution_channel[test_hb$distribution_channel == "TA/TO"] <- 4
test_hb$distribution_channel[test_hb$distribution_channel == "Undefined"] <- 5
table(test_hb$distribution_channel)
test_hb$distribution_channel <- as.numeric(test_hb$distribution_channel)
```

```{r}
table(test_hb$is_repeated_guest)
test_hb$is_repeated_guest <- as.numeric(test_hb$is_repeated_guest)
table(test_hb$is_repeated_guest)
```

```{r}
table(test_hb$deposit_type)
```


```{r}
test_hb$deposit_type[test_hb$deposit_type == "No Deposit"] <- 1
test_hb$deposit_type[test_hb$deposit_type == "Non Refund"] <- 2
test_hb$deposit_type[test_hb$deposit_type == "Refundable"] <- 3
table(test_hb$deposit_type)
test_hb$deposit_type <- as.numeric(test_hb$deposit_type)
```

```{r}
table(test_hb$customer_type)
```


```{r}
test_hb$customer_type[test_hb$customer_type == "Contract"] <- 1
test_hb$customer_type[test_hb$customer_type == "Group"] <- 2
test_hb$customer_type[test_hb$customer_type == "Transient"] <- 3
test_hb$customer_type[test_hb$customer_type == "Transient-Party"] <- 4
table(test_hb$customer_type)
test_hb$customer_type <- as.numeric(test_hb$customer_type)
```

```{r}
str(test_hb)
```



# Modeling

## NN

```{r}
library(neuralnet)
library(visdat)
library(caTools)
```

```{r}
library(dplyr)
train_hb2 = sample_n(train_hb, 5000)
```


```{r}

y_train_hb2 = as.numeric(train_hb2[,2])
x_train_hb2 = as.numeric(as.matrix(train_hb2[,-c(2,14,20,21,24,25,31)]))
x_train_hb2 = matrix(as.numeric(x_train_hb2),ncol=24)
#
# y_test = as.numeric(XX_te[,11])
# x_test = as.numeric(as.matrix(XX_test[,2:10]))
# x_test = matrix(as.numeric(x_test),ncol=9)
#
df_train = data.frame(cbind(x_train_hb2,y_train_hb2))
#df_test = data.frame(cbind(x_test,y_test))
```

```{r}
# NN - first model: one hidden layer with 5 neurons
# 
set.seed(100)
nn1 = neuralnet(y_train_hb2~V1+V2+V3+V4+V5+V6+V7+V8+V9+V10+V11+V12+V13+
                  V14+V15+V16+V17+V18+V19+V20+V21+V22+V23+V24,
                data = df_train, hidden = 3, linear.output = F, rep = 5, stepmax=1e7)
summary(nn1)
# show the results
head(nn1$result.matrix)
```
```{r}
# see the errors in each replicate
nn1$result.matrix[1,]
# select the replicate which has attained the smallest error
imin_rep=which.min(nn1$result.matrix[1,])
imin_rep

# compute the predicted values

yhat = round(nn1$net.result[[imin_rep]])
head(yhat)
# compute the Confusion Matrix and the accuracy
ConfMat1_train=addmargins(table(y_train_hb2,yhat))
ConfMat1_train
accuracy1_train=(ConfMat1_train[1,1]+ConfMat1_train[2,2])/(nrow(train_hb))
accuracy1_train
#

```


```{r}
y_valid_hb = as.numeric(valid_hb[,2])
x_valid_hb = as.numeric(as.matrix(valid_hb[,c(1,3:25)]))
x_valid_hb = matrix(as.numeric(x_valid_hb),ncol=24)

df_valid = data.frame(cbind(x_valid_hb,y_valid_hb))

#df_valid = data.frame(cbind(x_valid_hb))

```


```{r}
# predicting Class using the neural network
 pr.nn1 <- predict(nn1,df_valid, rep=imin_rep)
 head(pr.nn1)
 yhat_test = round(pr.nn1)
 head(yhat_test)
 # compute the Confusion Matrix and the accuracy
 ConfMat1_test=addmargins(table(y_valid_hb,yhat_test))
 ConfMat1_test
 accuracy1_test=(ConfMat1_test[1,1]+ConfMat1_test[2,2])/nrow(df_valid)
 accuracy1_test
 accuracy1_train
```

```{r}
library("writexl")
df_valid$pred_class <- yhat_test
write_xlsx(df_valid,"C:\\Users\\HP 14\\Desktop\\UNICT\\2nd Semester\\prediction-validation.xlsx")
```



    
## logistic regression

```{r}
nL<-dim(train_hb)[1]
nL

nT<-dim(valid_hb)[1]
nT


glm.fit2_L=glm(is_canceled~hotel+lead_time+arrival_date_year+
              arrival_date_month+stays_in_weekend_nights+arrival_date_day_of_month+
              stays_in_week_nights+arrival_date_week_number+adults+children+
              babies+meal+market_segment+distribution_channel+is_repeated_guest+
              previous_cancellations+previous_bookings_not_canceled+
              booking_changes+deposit_type+days_in_waiting_list+customer_type+
              adr+total_of_special_requests+required_car_parking_spaces,
              data=train_hb, family=binomial)
summary(glm.fit2_L, digits=3)
glm.probs2_L=predict.glm(glm.fit2_L,type="response")
glm.pred2_L=rep("No",nL)
glm.pred2_L[glm.probs2_L>.5]="Yes" # assign "Up" for units with P(Y="Up")>0.5
table(glm.pred2_L,train_hb$is_canceled) # confusion matrix
confMat2_L<-addmargins(table(glm.pred2_L,train_hb$is_canceled))
confMat2_L
delta2_L=(confMat2_L[1,2]+confMat2_L[2,1])/nL*100
delta2_L
```



```{r}
# compute the model on the test set
glm.probs2_T=predict.glm(glm.fit2_L, newdata=valid_hb, type="response") # note the option newdata
glm.pred2_T=rep("No",nT)
glm.pred2_T[glm.probs2_T>.5]="Yes" # assign "Up" for units with P(Y="Up")>0.5
table(glm.pred2_T,valid_hb$is_canceled) # confusion matrix
confMat2_T<-addmargins(table(glm.pred2_T,valid_hb$is_canceled))
confMat2_T
delta2_T=(confMat2_T[1,2]+confMat2_T[2,1])/nT*100
delta2_T
```

## Random Forest


```{r}
# BAGGING AND RANDOM FOREST
#
# We apply bagging and random forests to the Boston data, using the
# randomForest package in R.
#
# Note that the exact results obtained here may depend on the version of R 
# and the version of the randomForest package installed on your computer. 
#
# Recall that bagging is simply a special case of a random forest with m = p. 
# Therefore, the randomForest() function can random be used to perform 
# both random forests and bagging.
 install.packages("randomForest")
#
library(randomForest)
set.seed(1)
bag.boston=randomForest(medv~.,data=Boston,subset=train,mtry=13,importance=TRUE)
bag.boston
# The argument mtry=13 indicates that all 13 predictors should be considered
# for each split of the trees in other words, that bagging should be done.
# 
# Run this bagged model on the test set
yhat.bag = predict(bag.boston,newdata=Boston[-train,])
boston.test=Boston[-train,"medv"]
plot(yhat.bag, boston.test, pch=20, col="red", cex=0.8)
abline(0,1)
test_set.MSE_bag=mean((yhat.bag-boston.test)^2)
test_set.MSE_bag
sqrt(test_set.MSE_bag)
# The test set MSE associated with the bagged regression tree is 
# test_set.MSE(value), smaller than the value obtained using an 
# optimally-pruned single tree.
# 
# We could change the number of trees grown by randomForest() using the ntree argument
# (default value: ntree=500):

```

```{r}
library(randomForest)
set.seed(1)
bag.boston=randomForest(is_canceled~hotel+lead_time+arrival_date_year+
              arrival_date_month+stays_in_weekend_nights+arrival_date_day_of_month+
              stays_in_week_nights+arrival_date_week_number+adults+children+
              babies+meal+market_segment+distribution_channel+is_repeated_guest+
              previous_cancellations+previous_bookings_not_canceled+
              booking_changes+deposit_type+days_in_waiting_list+customer_type+
              adr+total_of_special_requests+required_car_parking_spaces,
              data=train_hb, mtry=24, ntree=10)
bag.boston
```

```{r}
yhat.bag = predict(bag.boston,newdata=valid_hb)

test_set.MSE_bag2=mean((yhat.bag-as.matrix(valid_hb[2]))^2)
test_set.MSE_bag2
sqrt(test_set.MSE_bag2)
```

```{r}
yhat_test = round(yhat.bag)
 head(yhat_test)
 # compute the Confusion Matrix and the accuracy
 ConfMat1_test=addmargins(table(y_valid_hb,yhat_test))
 ConfMat1_test
 accuracy1_test=(ConfMat1_test[1,1]+ConfMat1_test[2,2])/nrow(valid_hb)
 accuracy1_test
```


# We get essentially the same value as before
#
# Growing a random forest proceeds in exactly the same way, except that
# we use a smaller value of the mtry argument. By default, randomForest()
# uses p/3 variables when building a random forest of regression trees, and
# sqrt{p} variables when building a random forest of classification trees. Here we
# use mtry = 6.
#
# importance is a logical value to assess the importance of the variables
#
set.seed(1)
rf.boston=randomForest(medv~.,data=Boston,subset=train,mtry=6,importance=TRUE)
yhat.rf = predict(rf.boston,newdata=Boston[-train,])
test_set.MSE_rf=mean((yhat.rf-boston.test)^2)
test_set.MSE_rf
sqrt(test_set.MSE_rf)
sqrt(test_set.MSE_bag)
sqrt(test_set.MSE_bag2)
# random forests yielded an improvement over bagging in this case.
#
# Using the importance() function, we can view the importance of each variable.
importance(rf.boston)
# Two measures of variable importance are reported. 
#
# %IncMSE is based upon the mean decrease of accuracy in predictions 
# on the out of bag samples when a given variable is excluded from the model. 
#
# IncNodePurity is a measure of the total decrease in node impurity 
# that results from splits over that variable, averaged over all trees. 
#
# In the case of regression trees, the node impurity is measured by the training RSS, 
# and for classification trees by the deviance.
varImpPlot(rf.boston, col="blue", main="Boston data")
# The results indicate that across all of the trees considered in the random forest, 
# the wealth level of the community (lstat) and the house size (rm) are by far 
# the two most important variables.
#
```













































































































































































































































































































































































































































































































































































































































































































































































































































































































































































































































































































